---
output: github_document
---


<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "README-"
)

```

# oddwater

We propose a framework to detect anomalies in water quality sensor data.


This package is still under development and this repository contains a
development version of the R package *oddwater*.

## Installation

You can install oddwater from github with:

``` r
# install.packages("devtools")
devtools::install_github("pridiltal/oddwater")
```

## Example

### Data Preprocessing. These outliers were confirmed by DES
```{r preprocess, warning=FALSE, message=FALSE}
library(tidyverse)
library(oddwater)
#Label data
# Outliers confirmed by DES
outliers <- c("22-03-17 10:00", "05-04-17 7:20", "13-06-17 4:50", "03-11-17 7:50", "26-07-17 16:00")
# Neighbours of those outliers
neighbours <- c("22-03-17 11:10", "05-04-17 8:50", "13-06-17 6:20", "26-07-17 15:00", "03-11-17 9:20")
# label points as outliers or typical
sandy_creek$type <- ifelse( (sandy_creek$Timestamp  %in% outliers ), "outlier", "typical")
neighbour <- which(sandy_creek$Timestamp  %in% neighbours )
sandy_creek$type[neighbour] <- "neighbour"

```


### Transform data
```{r trans , dependson='preprocess'}
data <- sandy_creek[,c("Timestamp", "Lsonde_Cond_uscm", "Lsonde_Turb_NTU",      
                       "Lsonde_Level_m")] %>% drop_na()
trans_data <-oddwater::transform_data(data)
head(trans_data)
```

### Visualize time gap

`sandy_creek` data is an irregular time series

```{r time_gap, dependson = 'trans'}
trans_data <- trans_data %>% drop_na()
p <- ggplot(trans_data, aes(x= Timestamp, y= time))+
      geom_line()+
      ylab("time (minutes)")
print(p)
```

### Visualize original data
```{r vis_orig , dependson='trans', eval= FALSE }
plot_var <- dplyr::left_join(trans_data, sandy_creek, by = "Timestamp")

plot_var <- trans_data[, c("Timestamp", "Lsonde_Cond_uscm",   
                           "Lsonde_Turb_NTU", "Lsonde_Level_m" )] %>% drop_na()
oddwater::plot_series(plot_var, title = "original series") 
oddwater::plot_pairs(plot_var)

```


### Visualize transformed data
```{r vis_trans , dependson='trans', eval= FALSE}
plot_var <- trans_data[, c("Timestamp", "type", "neg_der_log_bounded_turb",
                           "pos_der_log_bounded_cond", "neg_der_log_bounded_level")] %>% drop_na()
colnames(plot_var) <- abbreviate(colnames(plot_var), minlength=20)
oddwater::plot_series(plot_var, title = "Non linear transformation - one side log derivatives (bounded")
oddwater::plot_pairs(plot_var)
``` 


### Compute performance metrics

```{r pm}
# Generate a toy dataset
true_labels <- c("out", "out", "normal", "out", "normal", "normal", "normal", "normal", "normal", "normal")
output <-  c("out", "normal", "normal", "normal", "out", "out", "normal", "normal", "normal", "normal")
# Compute performance metrics 
out<- calc_performance_metrics(y_truth = true_labels, y_output = output, pos_label = "out", neg_label = "normal")
```

Compute Means square Error
```{r eval = FALSE}
clac_MSE(true_y, pred_y)
```


## Further Details

```{r eval = FALSE}
?oddwater
```
