---
output: github_document
---

[![Project Status: WIP ? Initial development is in progress, but there has not yet been a stable, usable release suitable for the public.](https://www.repostatus.org/badges/latest/wip.svg)](https://www.repostatus.org/#wip)
[![Licence](https://img.shields.io/badge/licence-GPL--3-blue.svg)](https://www.gnu.org/licenses/gpl-3.0.en.html)
 
---
 
[![minimal R version](https://img.shields.io/badge/R%3E%3D-3.4.0-6666ff.svg)](https://cran.r-project.org/)
[![CRAN_Status_Badge](http://www.r-pkg.org/badges/version/oddwater)](https://cran.r-project.org/package=oddwater)
[![packageversion](https://img.shields.io/badge/Package%20version-0.5.0.9000-orange.svg?style=flat-square)](commits/master)
[![DOI](https://zenodo.org/badge/143960804.svg)](https://zenodo.org/badge/latestdoi/143960804)
 
---
 
[![Last-changedate](https://img.shields.io/badge/last%20change-`r gsub('-', '--', Sys.Date())`-yellowgreen.svg)](/commits/master)

[![Build Status](https://travis-ci.org/pridiltal/oddwater.svg?branch=master)](https://travis-ci.org/pridiltal/oddwater)

<!-- README.md is generated from README.Rmd. Please edit that file -->


```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-"
)

```

# oddwater

Outliers due to technical errors in water-quality data from in situ sensors can reduce data quality and have a direct impact on inference drawn from subsequent data analysis. Here, we proposed an automated framework that provides early detection of outliers in water-quality data from in situ sensors **caused by technical issues**. The framework was used first to identify the data features that differentiate outlying instances from typical behaviours. Then statistical transformations were applied to make the outlying instances stand out in transformed data space. Unsupervised outlier scoring techniques were then applied to the transformed data space and an approach based on extreme value theory was used to calculate a threshold for each potential outlier.

[In this work](https://www.monash.edu/__data/assets/pdf_file/0007/1645027/wp01-2019.pdf) we showed that our proposed framework, with carefully selected data transformation methods derived from data features, can greatly assist in increasing the
performance of a range of existing outlier detection algorithms.


This package is still under development and this repository contains a development version of the R package *oddwater*.

## Installation

You can install oddwater from github with:

``` r
# install.packages("devtools")
devtools::install_github("pridiltal/oddwater")
```


### Illustrative Example

#### Data obtained from in situ sensors at Sandy Creek.
```{r preprocess, warning=FALSE, message=FALSE}
library(tidyverse)
library(oddwater)
data("data_sandy_anom")
head(data_sandy_anom)
```


### Shiny App

```{r shiny, warning=FALSE, message=FALSE, eval=FALSE}
oddwater::explore_data()
```
